Flexible Probabilistic Forecasting with Distributional Trees and Forests

### 2018-03-19 pkg/disttree/inst/MIP-Seminar/slides.Rnw
small corrections and shorter list of references
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2061 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 532f82ea394628d36ebf14f42e9777ad485840a2

\vspace{-4cm}
\visible<6->{
\textbf{Application/Motivation:}
\begin{itemize}
\item Prediction.
\item Analysis of dependencies and interactions.
\end{itemize}
}
\vspace{0.5cm}
\visible<7->{
\textbf{Two main approaches / ``cultures'':}
\begin{itemize}
\item Data modeling.
\item Algorithmic modeling.
\end{itemize}
}
\end{frame}

@@

\visible<2->{
\begin{minipage}{0.65\textwidth}
\textbf{Example:} 
\only<2-3>{Linear model}\only<4>{Generalized additive model}\only<5>{Generalized additive model\\ for location, scale, and shape}
\\
\only<2-3>{
$$\visible<3>{\mathbb{E}(Y|X) = } f(X; \alpha) = \alpha_0 + \alpha_1 \cdot X$$  
}
\only<4>{
$$\mathbb{E}(Y|X) = f(X; \alpha) = \alpha + s(X)$$  
}
\only<5>{
\vspace{-0.8cm}
\begin{align*}
\mu(X) = \alpha_{\mu} + s_{\mu}(X)\\  
\sigma(X) = \alpha_{\sigma} + s_{\sigma}(X)
\end{align*}
}
\end{minipage}
%\vspace{-0.6cm}
%\hfill
\begin{minipage}{0.3\textwidth}
<<motivation_GLM, echo=FALSE, results=hide>>=
nobs <- 200
## GLM
set.seed(7)
x <- c(1:nobs)/nobs
ytrue <- 1+x*1.5
y <- ytrue + rnorm(nobs,0,0.3)
@
\only<2-3>{
<<plot_motivation_GLM_single, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
@
}
<<motivation_GAM, echo=FALSE, results=hide>>=
## GAM
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3 
y <- ytrue + rnorm(nobs,0,0.3)
@
\only<4>{
<<plot_motivation_GAM, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
@
}
<<motivation_GAMLSS, echo=FALSE, results=hide>>=
## GAMLSS
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3
var <- exp(-(2*x)^2)/2
y <- ytrue + rnorm(nobs, 0, 0.1 + var)
@
\only<5>{
<<plot_motivation_GAMLSS, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x, y, xaxt = "n", yaxt = "n", ann = FALSE, type = "n")
polygon(c(x, rev(x)), c(ytrue + 0.1 + var, rev(ytrue - 0.1 - var)),
  col = pallight["forest"], border = "transparent")
lines(x, ytrue, col = pal["forest"], lwd=7)
points(x, y, col = "slategray", pch = 19)
box(lwd = 5)
@
}
\end{minipage}
}
\end{frame}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\SweaveOpts{eval=FALSE}

@@

\vspace{-0.2cm}

\footnotesize
Hothorn T, Hornik K, Zeileis A (2006).
 \dquote{Unbiased Recursive Partitioning: A Conditional Inference Framework.}
 \textit{Journal of Computational and Graphical Statistics},
 \textbf{15}(3), 651--674.
 \doi{10.1198/106186006X133933}

Zeileis A, Hothorn T, Hornik K (2008).
 \dquote{Model-Based Recursive Partitioning.}
  \textit{Journal of Computational and Graphical Statistics},
  \textbf{17}(2), 492--514.
  \doi{10.1198/106186008X319331}

@@

@@

Gneiting T, Raftery A E, Westveld III A H, Goldman T (2005).
  \dquote{Calibrated Probabilistic Forecasting using Ensemble Model Output Statistics and Minimum CRPS Estimation}
  \textit{Monthly Weather Review}, 
  \textbf{133}(5), 1098--1118.
  \doi{10.1175/MWR2904.1}

Hofner B, Mayr A, Schmid M (2016).
  \dquote{\textbf{gamboostLSS}: An \textsf{R} Package for Model Building and Variable Selection in the {GAMLSS} Framework}
  \textit{Journal of Statistical Software},
  \textbf{74}(1), 1--31.
  \doi{10.18637/jss.v074.i01}

@@

\begin{frame}
\frametitle{References (2)}

\vspace{-0.2cm}

% Reference for the observation data set
BMLFUW (2016).
  \dquote{{B}undesministerium f\"ur {L}and und {F}orstwirtschaft, {U}mwelt und {W}asserwirtschaft ({BMLFUW}), {A}bteilung {IV/4} -- {W}asserhaushalt}
  \text{Available at \url{http://ehyd.gv.at}}, 
  \text{Accessed: 2016--02--29}.

% Reference for the GFS reforecast version 2 data set 
Hamill T M, Bates G T, Whitaker J S, Murray D R, Fiorino M, Galarneau T J Jr., Zhu Y, Lapenta W (2013).
  \dquote{{NOAA}'s Second-Generation Global Medium-Range Ensemble Reforecast Dataset}
  \textit{Bulletin of the American Meteorological Society}, 
  \textbf{94}(10), 1553--1565.
  \doi{10.1175/BAMS-D-12-00014.1}

@@

Messner J W,  Mayr G J, Zeileis A (2016).
  \dquote{Heteroscedastic Censored and Truncated Regression with \textbf{crch}}
  \textit{The \textsf{R} Journal}, 
  \textbf{8}(1), 173--181.\\
  \scriptsize
  \url{https://journal.R-project.org/archive/2016-1/messner-mayr-zeileis.pdf}
  \footnotesize

Stauffer R, Mayr G J, Messner J W, Umlauf N, Zeileis A (2017).
  \dquote{Spatio-Temporal Precipitation Climatology Over Complex Terrain Using a Censored Additive Regression Model}
  \textit{International Journal of Climatology}, 
  \textbf{37}(7), 3264--3275.
  \doi{10.1002/joc.4913}

Gebetsberger M, Messner J W, Mayr G J, Zeileis A (2017).
  \dquote{Fine-Tuning Non-Homogeneous Regression for Probabilistic Precipitation Forecasts: Unanimous Predictions, Heavy Tails, and Link Functions}
  \textit{Monthly Weather Review}, 
  \text{Forthcoming}.
  \doi{10.1175/MWR-D-16-0388.1}

@@

>>>>

\vspace{-4cm}
\visible<6->{
\textbf{Application/Motivation:}
\begin{itemize}
\item Prediction.
\item Caption of dependencies and interactions.
\end{itemize}
}
\vspace{0.5cm}
\visible<7->{
\textbf{Two main approaches / ``cultures'':}
\begin{itemize}
\item Data modeling.
\item Algorithmic modeling.
\end{itemize}
}
\end{frame}

@@

\visible<2->{
\begin{minipage}{0.65\textwidth}
\textbf{Example:} 
\only<2-3>{Linear model}\only<4>{Generalized additive model}\only<5-6>{Generalized additive model\\ for location, scale, and shape}
\\
\only<2-3>{
$$\visible<3>{\mathbb{E}(Y|X) = } f(X; \alpha) = \alpha_0 + \alpha_1 \cdot X$$  
}
\only<4>{
$$\mathbb{E}(Y|X) = f(X; \alpha) = \alpha + s(X)$$  
}
\only<5-6>{
\vspace{-0.8cm}
\begin{align*}
\visible<6>{\mathbb{E}(Y|X) = }\mu(X) = \alpha_{\mu} + s_{\mu}(X)\\  
\visible<6>{\sqrt{Var(Y|X)} = }\sigma(X) = \alpha_{\sigma} + s_{\sigma}(X)
\end{align*}
}
\end{minipage}
%\vspace{-0.6cm}
%\hfill
\begin{minipage}{0.3\textwidth}
<<motivation_GLM, echo=FALSE, results=hide>>=
nobs <- 200
## GLM
set.seed(7)
x <- c(1:nobs)/nobs
ytrue <- 1+x*1.5
y <- ytrue + rnorm(nobs,0,0.3)
@
\only<2-3>{
<<plot_motivation_GLM_single, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
@
}
<<motivation_GAM, echo=FALSE, results=hide>>=
## GAM
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3 
y <- ytrue + rnorm(nobs,0,0.3)
@
\only<4>{
<<plot_motivation_GAM, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
@
}
<<motivation_GAMLSS, echo=FALSE, results=hide>>=
## GAMLSS
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3
var <- exp(-(2*x)^2)/2
y <- ytrue + rnorm(nobs, 0, 0.1 + var)
@
\only<5-6>{
<<plot_motivation_GAMLSS, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x, y, xaxt = "n", yaxt = "n", ann = FALSE, type = "n")
polygon(c(x, rev(x)), c(ytrue + 0.1 + var, rev(ytrue - 0.1 - var)),
  col = pallight["forest"], border = "transparent")
lines(x, ytrue, col = pal["forest"], lwd=7)
points(x, y, col = "slategray", pch = 19)
box(lwd = 5)
@
}
\end{minipage}
}
\end{frame}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\SweaveOpts{eval=TRUE}

@@

\vspace{0.5cm}
%\vspace{-0.2cm}

%Hothorn T, Hornik K, Zeileis A (2006).
% \dquote{Unbiased Recursive Partitioning: A Conditional Inference Framework.}
% \textit{Journal of Computational and Graphical Statistics},
% \textbf{15}(3), 651--674.
% \doi{10.1198/106186006X133933}

%Zeileis A, Hothorn T, Hornik K (2008).
% \dquote{Model-Based Recursive Partitioning.}
%  \textit{Journal of Computational and Graphical Statistics},
%  \textbf{17}(2), 492--514.
%  \doi{10.1198/106186008X319331}

Schlosser L, Hothorn T, Stauffer R, Zeileis A (2018).
  \dquote{Distributional Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain.}
  \emph{Manuscript under preparation}.

\bigskip

@@

\bigskip

@@

%Gneiting T, Raftery A E, Westveld III A H, Goldman T (2005).
%  \dquote{Calibrated Probabilistic Forecasting using Ensemble Model Output Statistics and Minimum CRPS Estimation}
%  \textit{Monthly Weather Review}, 
%  \textbf{133}(5), 1098--1118.
%  \doi{10.1175/MWR2904.1}

%Hofner B, Mayr A, Schmid M (2016).
%  \dquote{\textbf{gamboostLSS}: An \textsf{R} Package for Model Building and Variable Selection in the {GAMLSS} Framework}
%  \textit{Journal of Statistical Software},
%  \textbf{74}(1), 1--31.
%  \doi{10.18637/jss.v074.i01}

@@

%\begin{frame}
%\frametitle{References (2)}

%\vspace{-0.2cm}

%\footnotesize

%% Reference for the observation data set
%BMLFUW (2016).
%  \dquote{{B}undesministerium f\"ur {L}and und {F}orstwirtschaft, {U}mwelt und {W}asserwirtschaft ({BMLFUW}), {A}bteilung {IV/4} -- {W}asserhaushalt}
%  \text{Available at \url{http://ehyd.gv.at}}, 
%  \text{Accessed: 2016--02--29}.

% Reference for the GFS reforecast version 2 data set 
%Hamill T M, Bates G T, Whitaker J S, Murray D R, Fiorino M, Galarneau T J Jr., Zhu Y, Lapenta W (2013).
%  \dquote{{NOAA}'s Second-Generation Global Medium-Range Ensemble Reforecast Dataset}
%  \textit{Bulletin of the American Meteorological Society}, 
%  \textbf{94}(10), 1553--1565.
%  \doi{10.1175/BAMS-D-12-00014.1}

@@

%Messner J W,  Mayr G J, Zeileis A (2016).
%  \dquote{Heteroscedastic Censored and Truncated Regression with \textbf{crch}}
%  \textit{The \textsf{R} Journal}, 
%  \textbf{8}(1), 173--181.\\
%  \scriptsize
%  \url{https://journal.R-project.org/archive/2016-1/messner-mayr-zeileis.pdf}
%  \footnotesize

%Stauffer R, Mayr G J, Messner J W, Umlauf N, Zeileis A (2017).
%  \dquote{Spatio-Temporal Precipitation Climatology Over Complex Terrain Using a Censored Additive Regression Model}
%  \textit{International Journal of Climatology}, 
%  \textbf{37}(7), 3264--3275.
%  \doi{10.1002/joc.4913}

%Gebetsberger M, Messner J W, Mayr G J, Zeileis A (2017).
%  \dquote{Fine-Tuning Non-Homogeneous Regression for Probabilistic Precipitation Forecasts: Unanimous Predictions, Heavy Tails, and Link %Functions}
%  \textit{Monthly Weather Review}, 
%  \text{Forthcoming}.
%  \doi{10.1175/MWR-D-16-0388.1}

@@

%\end{frame}
